Arrival Management with Open Data

### 2023-04-04 paper.qmd
initial stab at background section

---
# template taken from https://github.com/gael-close/ieeetran
#----------------------------------------------------------
title: Arrival Management with Open Data
format:
  ieeetran-pdf:
    keep-tex: true

@@

## header 2

@@

<!-- what did this paper show -->

This paper aimed at exploring a data-driven approach to measuring arrival management based on open data.

<!-- strength and weaknesses -->

<!-- application -->

<!-- further research -->

<!-- final appraisal -->

@@

The script to download the weekly global datasets are included.
The cleaned trajectory data is stored at: \<\<tbd\>\>.

@@

The authors thank the comments by XX YY ZZ.

The views expressed are the authors' own and do not represent a policy or position of EUROCONTROL <!-- and co-authors organisations -->.

>>>>

---
# template taken from https://github.com/gael-close/ieeetran
#----------------------------------------------------------
title: Arrival Management with Open Data
format:
  ieeetran-pdf:
    keep-tex: true
  #docx: default

@@

Trajectory-based operations (TBO) are a core element of the ongoing transformation programs for air navigation services (e.g. Europe/SESAR, US/Next Gen, Japan/CARATS).
On the conceptual level TBO describe a move from clearance-based operations to a highly predictable and negotiated trajectory execution between airspace users and air traffic control.
The TBO concept entails the successive refinement (i.e. negotiation) between the parties starting from an inital planned trajectory through operations and the actual execution of flight.
TBO builds on higher levels of information exchange between planning and flow management systems and the airborne and ground systems during flight operations. @fig-TBO-concept shows the overall timeline and evolution of a trajectory.

```{r}
#| label: fig-TBO-concept
#| fig-cap: Trajectory-based operations - conceptual perspective
#| out-width: 45%  # 2 column IEEE
knitr::include_graphics("./figures/TBO-concept.png")
```

Conceptual approach := technology agnostic
CONOPS may entail technological enablers and/or advanced procedures
Performance objectives observed on concept functional level

@@

<!-- what did this paper show -->
Ambitious political targets for 2050 have been formulated to curb the contribution of air transportation to climate change and work towards climate neutrality.
Policy and industry action revolve around a variety of measures. 
Within this context - and in light of the required lead time for a global pickup on sustainable aviation fuels, the introduction of novel airframe designs, or propulsion technology - increasing the operational efficiency is a declared now-term measure.
There is a substantial benefit pool for the wider arrival phase covering the airspace within 200NM of the destination airport.
Trajectory-based operations promise higher levels of predictability and reduced inefficiencies. 
The benefits of reducing spacing and flow synchronisation measures in the terminal airspace are therefore dependent on the capability of balancing out operational variations from delivery of the aircraft to the wider arrival airspace.

This proof-of-concept paper aims at operationalising a performance measure to characterise arrival management techniques, but also provide a basis to discuss the handover from the enroute phase to the wider arrival airspace.

This paper aimed at exploring a data-driven approach to measuring operational efficiency during the arrival phase. 
In light of the ongoing discussions about the climate impact of air transportation and the contr
Arrival management based on open data.

<!-- strength and weaknesses -->
The proposed approach builds on the generalisation of a well-understood key performance metric for the wider arrival airspace.
It follows a data-driven approach that is independent of detailed local context modelling.
For example, local airspace restrictions or procedures are embedded in the temporal-spatial framework.
Measuring the spacing deviation between interacting arrival flights provides a basis for the analysis of approach concepts that can help to streamline the arrival sequencing and contribute to the further reduction of unnecessary fuel burn and emissions.

This proof-of-concept addresses the aspect of transparency by making use of openly available air transport trajectory data and open (analytical) software. The approach is readily transferrable to other regions.

<!-- application -->
The goal of this work was the initial development of an operational performance measure.
Further validation is required at a larger set of European (and global airports).
This will allow for the refinement of the parameterization of the approach.
It is planned to roll-out the approach under the umbrella of the EUROCONTROL Performance Review System.
This will provide European stakeholders with a means to address enhancement of arrival operations, and may help to build a catalogue of arrival management techniques that support higher levels of efficiency.

<!-- further research -->
This paper demonstrated feasibility of the approach. While the proposed performance monitoring of the spacing deviation envelope provides a basis for identifying inefficiencies during the arrival flow, there is a need to account also for factors driving and impacting the observed performance. 
The temporal assessment of spacing deviation

* weather, wind
* runway system configuration, adjacent aerodromes/level capping

On the algorithmic level, future work should address the impact of the geospatial dimensions of cells and the method of their definition.

This paper applied a observational approach to determining the reference times per cell. 
This will pose challenges for longer time horizons as the continual update and processing of arrival trajectories would require substantial resources. 
A possibility is the use of data structures to support the continual online accumulation of statistics without the need to store the complete historic dataset or processed artefacts.

<!-- final appraisal -->
The results demonstrate the general feasibility of the approach.
The proposed approach offers a measure to start the dialogue between operational planners and stakeholders to identify benefit pools during the arrival phase.

@@

<!--
The script to download the weekly global datasets are included.
The cleaned trajectory data is stored at: \<\<tbd\>\>.
-->

@@

The authors thank the comments by colleagues from the Performance Section of the Department of Airspace Control (DECEA) Brazil and the Performance Review Unit of EUROCONTROL. 
This study would have not been feasible without the generous support of Opensky Network and the many community members collecting and feeding open ADSB data.

The views expressed are the authors' own and do not represent a policy or position of DECEA or EUROCONTROL <!-- and co-authors organisations -->.

### 2023-04-04 paper.tex
initial stab at background section

\hypertarget{header-2}{%
\subsection{header 2}\label{header-2}}

@@

\begin{equation}\protect\hypertarget{eq-spacing-deviation}{}{\begin{aligned}
spacing deviation (t) =  ref{\_}time(trailer_{(t)}) - \\ref{\_}time(leader_{(t - s)})
\end{aligned}}\label{eq-spacing-deviation}\end{equation}

@@

This paper aimed at exploring a data-driven approach to measuring
arrival management based on open data.

@@

The script to download the weekly global datasets are included. The
cleaned trajectory data is stored at:
\textless\textless tbd\textgreater\textgreater.

@@

The authors thank the comments by XX YY ZZ.

The views expressed are the authors' own and do not represent a policy
or position of EUROCONTROL .

>>>>

Trajectory-based operations (TBO) are a core element of the ongoing
transformation programmes (e.g.~Europe/SESAR, US/Next Gen,
Japan/CARATS). On the conceptual level TBO describe a move from
clearance-based operations to a highly predictable and negotiated
trajectory execution between airspace users and air traffic control. The
TBO concept entails the successive refinement (i.e.~negotiation) between
the parties starting from an inital planned trajectory through
operations and the actual execution of flight. TBO builds on higher
levels of information exchange between planning and flow management
systems and the airborne and ground systems during flight operations.
Figure~\ref{fig-TBO-concept} shows the overall timeline and evolution of
a trajectory.

{\centering \includegraphics[width=0.45\textwidth,height=\textheight]{./figures/TBO-concept.png}

\caption{\label{fig-TBO-concept}Trajectory-based operations - conceptual
perspective}

Conceptual approach := technology agnostic CONOPS may entail
technological enablers and/or advanced procedures Performance objectives
observed on concept functional level

@@

\begin{equation}\protect\hypertarget{eq-spacing-deviation}{}{\begin{aligned}
spacing \text{ } deviation (t) =  ref{\_}time(trailer_{(t)}) - \\ref{\_}time(leader_{(t - s)})
\end{aligned}}\label{eq-spacing-deviation}\end{equation}

@@

Ambitious political targets for 2050 have been formulated to curb the
contribution of air transportation to climate change and work towards
climate neutrality. Policy and industry action revolve around a variety
of measures. Within this context - and in light of the required lead
time for a global pickup on sustainable aviation fuels, the introduction
of novel airframe designs, or propulsion technology - increasing the
operational efficiency is a declared now-term measure. There is a
substantial benefit pool for the wider arrival phase covering the
airspace within 200NM of the destination airport. Trajectory-based
operations promise higher levels of predictability and reduced
inefficiencies. The benefits of reducing spacing and flow
synchronisation measures in the terminal airspace are therefore
dependent on the capability of balancing out operational variations from
delivery of the aircraft to the wider arrival airspace.

This proof-of-concept paper aims at operationalising a performance
measure to characterise arrival management techniques, but also provide
a basis to discuss the handover from the enroute phase to the wider
arrival airspace.

This paper aimed at exploring a data-driven approach to measuring
operational efficiency during the arrival phase. In light of the ongoing
discussions about the climate impact of air transportation and the contr
Arrival management based on open data.

The proposed approach builds on the generalisation of a well-understood
key performance metric for the wider arrival airspace. It follows a
data-driven approach that is independent of detailed local context
modelling. For example, local airspace restrictions or procedures are
embedded in the temporal-spatial framework. Measuring the spacing
deviation between interacting arrival flights provides a basis for the
analysis of approach concepts that can help to streamline the arrival
sequencing and contribute to the further reduction of unnecessary fuel
burn and emissions.

This proof-of-concept addresses the aspect of transparency by making use
of openly available air transport trajectory data and open (analytical)
software. The approach is readily transferrable to other regions.

The goal of this work was the initial development of an operational
performance measure. Further validation is required at a larger set of
European (and global airports). This will allow for the refinement of
the parameterization of the approach. It is planned to roll-out the
approach under the umbrella of the EUROCONTROL Performance Review
System. This will provide European stakeholders with a means to address
enhancement of arrival operations, and may help to build a catalogue of
arrival management techniques that support higher levels of efficiency.

This paper demonstrated feasibility of the approach. While the proposed
performance monitoring of the spacing deviation envelope provides a
basis for identifying inefficiencies during the arrival flow, there is a
need to account also for factors driving and impacting the observed
performance. The temporal assessment of spacing deviation

\begin{itemize}
\tightlist
\item
  weather, wind
\item
  runway system configuration, adjacent aerodromes/level capping
\end{itemize}

On the algorithmic level, future work should address the impact of the
geospatial dimensions of cells and the method of their definition.

This paper applied a observational approach to determining the reference
times per cell. This will pose challenges for longer time horizons as
the continual update and processing of arrival trajectories would
require substantial resources. A possibility is the use of data
structures to support the continual online accumulation of statistics
without the need to store the complete historic dataset or processed
artefacts.

The results demonstrate the general feasibility of the approach. The
proposed approach offers a measure to start the dialogue between
operational planners and stakeholders to identify benefit pools during
the arrival phase.

@@

@@

The authors thank the comments by colleagues from the Performance
Section of the Department of Airspace Control (DECEA) Brazil and the
Performance Review Unit of EUROCONTROL. This study would have not been
feasible without the generous support of Opensky Network and the many
community members collecting and feeding open ADSB data.

The views expressed are the authors' own and do not represent a policy
or position of DECEA or EUROCONTROL .
